---
title: "Eating Out"
author: "Tyler Hamblin"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`"    
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## There were two years in which the General Social Survey asked about the frequency of eating out. The variable name you are looking for is "eatout".
### Find those years and for the first year the question was asked download that years data in a stata format - its name will be "GSSxxxx.dta" where the xxxx is replaced by the year.  For example if the year was 1983 its name would be GSS1983.dta. Your assignment is to download the GSS yearly dataset for the first year the question was asked and then  write up the answer to the question by editing the  code block in line 31 to eval=TRUE and finish the job. 



$H_0$ There is no difference among the various races with frequency of eating out  
$H_A$ There is a statistical difference between at least two races with the frequency of eating out  


```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
```
```{r, eval=TRUE, warning=FALSE}
#GSS <-read_dta(file="GSS1996.dta")
GSS <- readstata13::read.dta13(file="GSS1996.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,age,eatout)
mydata <- na.omit(mydata)
```


## Graphical Data 

```{r}
barchartGC(~race + eatout,data=mydata)
barchartGC(~race + eatout,data=mydata, type="percent")

```

When looking at the non-percentage data it appears that there is a disproportionate amount of white people eating out when compared to the black and other categories. However, when comparing the percentage of people responding "yes" or "no" to the amount of people from each people group included in the survey, we get percentage based statistics on the second graph. Here we can see that the bar graphs are a lot closer in proportion with white people responding yes to eating out the most and black people having the highest rate of answering no.





## Numerical Data

```{r}
table1 <- xtabs(~race + eatout, data=mydata)
rowPerc(table1)
colPerc(table1)
```

The numerical results confirm what was predicted from the graphical data but more accurately we can now say that white people say yes the most at around 82%, which is 14% higher than black and 5% higher than other, and black people say no the most at around 31%, which is 14% higher than white people and 9% higher than other. But again we can see that the percentage of people answering yes or no to this survey is significantly primarily white people (82% of people in entire yes category were white and 71% of people in entire no category). 



## Inferential Data 

```{r}
fisherXY(mydata$race,mydata$eatout, printAll = TRUE)
```

The fisherXY test shows a slight variation in results than the confidence that the previous two tests suggested. There is a large p-value (0.2815) that suggests a large amount of room for error or the possibility that the null hypothesis could be true. In the 95% confidence interval we see that the perceived difference between the other and black categories may be explained by chance and a statistical difference in the data may not be the correct explanation. The black and white races however do have some statistical variation backed by the inferential test that is based on the p-value 0.0003864 and the confidence interval actually allows us to throw out the null hypothesis in this instance. The white and other categories were also fairly accurately depicted in the graphical and numerical data as we predicted that they were both fairly similar to each other when responding yes or no to eating out. The high p-value 0.3543 confirms that there is not enough variation for race to accurately distinguish these two categories for this question.  


## Discussion and Conclusion 

The process of importing the GSS data set directly off of the NORC website was not the most intuitive process because of the recent change that STATA files are read and interpreted by the r-program. Our graphical and numerical data sections provided us with little room to be confident in the slight differences between the race categories that could be seen. The massive sample size of the white people surveyed for this data is most likely responsible for this confusion and the contradicting conclusion of some of these results. The lack of consistency from the black and other people groups due to their comparably small sample sizes does not allow us to definitively say that race is a factor for eating out. 

